---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# qvirus

<!-- badges: start -->
<!-- badges: end -->

The goal of qvirus is to provide resources, tutorials, and code snippets dedicated to exploring the intersection of quantum computing and artificial intelligence (AI) in the context of HIV drug design.

## Installation

You can install the development version of qvirus from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("juanv66x/qvirus")
```

## Example

This is a basic example which shows you how to solve a common problem in quantum computing.

```{r example}
library(qvirus)
## basic example code
```

Creating a Hadamard gate: `H(1)`.

```{r hada, echo = FALSE}
library(qsimulatR)
H(1)
```

Applying a gate to a quantum state.

```{r gate}
library(qsimulatR)
H(1)*six_state(1)[[1]]
```

